### 2016-02-19 paper.tex
Add another comment about importance to MHK

A selection of measured performance data in the literature and its usage in
numerical model validation is presented in Table~\ref{tab:validation-data}.
Turbine diameter Reynolds numbers spanned from small laboratory scale ($Re_D
\sim 10^5$) all the way to full scale ($Re_D \sim 10^7$). Individual blade
forces were only measured in two of the experiments---Strickland \emph{et al.}
\cite{Strickland1981} and Laneville and Vitecoq \cite{Laneville1986}. There has
been nearly equal attention given to the eggbeater-shaped Darrieus rotor and the
straight-bladed H-rotor. The two additional rotors studied were the large scale
tapered H-rotor VAWT 850 \cite{Mays1990} and the helical Quiet Revolution QR5
measured by Penna \cite{Penna2008}, the exact geometry for which is unknown.

@@

In general, there have been more experiments done with rotors of low $c/R$,
which makes them easier to model, since unsteady dynamic effects are less
influential on the overall performance \cite{Strickland1981}. This is apparent
when examining the effectiveness of numerical models that rely on static foil
coefficient input data, e.g., streamtube and vortex models, which seem to be
most applicable for $c/R \leq 0.1$. For example, Bedon \emph{et al.}
\cite{Bedon2014} used a double multiple streamtube momentum model without
dynamic stall corrections to evaluate the effectiveness of various foil
coefficient databases against the Sandia Darrieus turbine experimental data.
Despite using such a simple model, performance predictions were quite accurate
in most conditions except at low tip speed ratio for the 2 m turbine, which had
the highest $c/R$ of all the Sandia rotors, making the dynamic effects more
important. This point highlights the need for more validation data for higher
solidity rotors to ensure numerical models are robust enough to explore unique
cross-flow turbine designs.

>>>>

A selection of measured performance data in the literature and its usage in
numerical model validation is presented in Table~\ref{tab:validation-data}.
Turbine diameter Reynolds numbers spanned from small laboratory scale ($Re_D
\sim 10^5$) all the way to full scale ($Re_D \sim 10^7$). Individual blade
forces were only measured in two of the experiments---Strickland \emph{et al.}
\cite{Strickland1981} and Laneville and Vitecoq \cite{Laneville1986}. There has
been nearly equal attention given to the eggbeater-shaped Darrieus rotor and the
straight-bladed H-rotor. The two additional rotors studied were the large scale
tapered H-rotor VAWT 850 \cite{Mays1990} and the helical Quiet Revolution QR5
measured by Penna \cite{Penna2008}, the exact geometrical details for which are
not available.

\todo[inline]{Remove QR5 data, since without geometry, no one else can use the
    data for validation?}

@@

In general, there have been more experiments done with rotors of low $c/R$,
which makes them easier to model, since unsteady dynamic effects are less
influential on the overall performance \cite{Strickland1981}. This is apparent
when examining the effectiveness of numerical models that rely on static foil
coefficient input data, e.g., streamtube and vortex models, which are most
applicable for $c/R \leq 0.1$. For example, Bedon \emph{et al.}~\cite{Bedon2014}
used a double multiple streamtube momentum model without dynamic stall
corrections to evaluate the effectiveness of various foil coefficient databases
against the Sandia Darrieus turbine experimental data. Despite using such a
simple model, performance predictions were quite accurate in most conditions
except at low tip speed ratio for the 2 m turbine, which had the highest $c/R$
of all the Sandia rotors, making the dynamic effects more important. This point
highlights the need for more validation data for higher solidity rotors to
ensure numerical models are robust enough to explore unique cross-flow turbine
designs, especially as the MHK concepts mature.
